---
title: "Reading data from web"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(sqldf)
library(httr)
library(jsonlite)
```


##################################
# READING CSV FILES FROM THE WEB #
##################################

#Add the URL inside the quotes

```{r}
url = "https://healthdata.gov/sites/default/files/estimated_inpatient_covid_20201201%201933.csv"
```


#Assign the csv data to a data frame
```{r}
covid_inpt = read.csv(url)
```


#Show the column names in the data frame
```{r}
names(covid_inpt)
```


#Show the first few rows of the data frame
```{r}
head(covid_inpt)
```


####################### SQL #######################

install.packages("sqldf")


##### To refer to variables with a period in the name, surround the variable name in [] #####

#QUESTION 1 - Are the earliest and most recent dates reported the same for all states/terrories?  
#If so, give the date of earliest and most recent reporting.  If not, list the different dates reported by state.
```{r q1_1}
report_date_count = sqldf(
  "WITH rprt_date AS 
       (SELECT state, MIN(collection_date) AS earliest_rprt_date, MAX(collection_date) AS latest_rprt_date
       FROM covid_inpt
       GROUP BY state)
    SELECT COUNT(DISTINCT(earliest_rprt_date)), COUNT(DISTINCT(latest_rprt_date))
    FROM rprt_date
  ") 
report_date_count


earliset_last_date = sqldf(
  "SELECT MIN(collection_date) AS earliest_rprt_date, MAX(collection_date) AS latest_rprt_date
   FROM covid_inpt
  "
)

earliset_last_date

#The earliest and most recent dates reported are the same for all states/terrories. 
#The earlist date is 2020-10-31 and the moset recent date is 2020-11-30. 
```



#QUESTION 2 - Which state/territory had the highest percentage of inpatient beds occupied by COVID-19
#patients on 11/28/2020?  What percentage of inpatient beds were occupied?

```{r}
inpatient_beds_occupied = sqldf(
  "SELECT state, [Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated] AS percentage
   FROM covid_inpt
   WHERE collection_date = '2020-11-28'
   GROUP BY state
   ORDER BY percentage DESC
   LIMIT 1
  "
)

inpatient_beds_occupied

#The highest percentage of inpatient beds occupied by COVID-19 patients on 11/28/2020 is 23.13% in North Dakota. 
```


#QUESTION 3 - Which date and state/territory had the highest ever percentage of inpatient beds occupied by COVID-19
#patients since recording began?

```{r}
highest_prctge = sqldf(
  "SELECT state, collection_date, [Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated] AS percentage
   FROM covid_inpt
   WHERE percentage IN 
      (SELECT MAX([Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated]) AS highest_percentage
       FROM covid_inpt) 
  "
)

# North Dakota on 2020-11-13 had the highest ever percentage (25.89%) of inpatient beds occupied by COVID-19 patients since recording began. 

```


#QUESTION 4 - What were the 3 worst days (by percentage) recorded for inpatient beds occupied by COVID-19 patients
#in New York State?  Report the date, number, and percentage of inpatient beds occupied.
```{r}

worst_days = sqldf(
  "SELECT collection_date, [Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated] AS number, [Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated] AS percentage
   FROM covid_inpt
   WHERE state = 'NY' AND
         percentage IN (
            SELECT [Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated]
            FROM covid_inpt
            WHERE state = 'NY'
            ORDER BY [Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated] DESC
            LIMIT 3)
  "
)

# The three worst days in NY state are 2020-11-28, 2020-11-29, and 2020-11-30. 
```


#QUESTION 5 - Which state/territory appears to have been most successful during the reporting period in
#keeping COVID-19 patients out of inpatient beds?  Base on the average percentage of inpatient beds
#utilized over the reporting period.

```{r}
best_state = sqldf(
  "SELECT state, AVG([Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated]) AS avg_pct
   FROM covid_inpt
   GROUP BY state
   HAVING avg_pct IN 
      (SELECT AVG([Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated]) AS min_pct
       FROM covid_inpt
       GROUP BY state
       ORDER BY min_pct
       LIMIT 1)
  "
)

# Vermont appears to have been most successful during the reporting period in keeping COVID-19 patients out of inpatient beds
```


#QUESTION 6 - Using the same query you wrote in #5, adjust it to show the top 10 most successful states
#during the reporting period.   
#Is New York State among the top 10? If yes, give its rank and average percentage.

```{r}
top_ten = sqldf(
  "WITH ten_states AS(
       SELECT state, AVG([Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated]) AS avg_pct
       FROM covid_inpt
       GROUP BY state
       HAVING avg_pct IN 
          (SELECT AVG([Percentage.of.Inpatient.Beds.Occupied.by.COVID.19.Patients.Estimated]) AS min_pct
           FROM covid_inpt
           GROUP BY state
           ORDER BY min_pct
           LIMIT 10)
       ORDER BY avg_pct)
   SELECT *, RANK() OVER(ORDER BY avg_pct) AS rank_pct
   FROM ten_states
  "
)

top_ten

# New York state is among the top 10. It ranks 8 with the average percentage of 6.373548
```



###################################################################
# PULLING JSON DATA FROM THE WEB USING HTTR AND JSONLITE PACKAGES #
###################################################################


install.packages("jsonlite")  #This package will help convert JSON data to a data frame

```{r}
library(jsonlite)
```


#Restrict this data only to ZIP code 10032 and use SQL to provide an insight about restaurant 
#violations in the campus neighborhood.

#Pull restaurant violations data for Washington Heights

```{r}
wahi <- GET("https://data.cityofnewyork.us/resource/43nn-pn8j.json", query = list("zipcode" = 10032))

new <- content(wahi, "text")
```



#Create data frame from JSON data

```{r}
wahi_df <- fromJSON(new) 
#NOTE: this will be limited to 1000 records due to throttling
```

#Show variable names for new data frame
colnames(wahi_df)


####################### SQL #######################

#QUESTION 7 - How many critical violations are reported in this sample of inspections?
```{r}
violation_count = sqldf(
  "SELECT COUNT(violation_code) AS vio_count, COUNT(DISTINCT(violation_code)) AS vio_type_count
   FROM wahi_df
  "
)

violation_count

#There are 980 violations in 58 categories have been reported in this sample. 

```


#QUESTION 8 - Give the name and address (building, street) of restaurant(s) with the highest 
#number of critical violations.  Account for possible ties in your results.
```{r}
highest_vio = sqldf(
  "SELECT dba, building, street, COUNT(violation_code) AS vio_count
   FROM wahi_df
   GROUP BY dba
   HAVING vio_count IN 
      (SELECT COUNT(violation_code) AS count
       FROM wahi_df
       GROUP BY dba
       ORDER BY count DESC
       LIMIT 1)
  "
)

highest_vio

#The restaurant with the highest number of violations is the BROTHER FISH MARKET, which located at 3845 Broadway and has 43 violations in total. 
```


#QUESTION 9 - Similarly to question 8, give the name and address (building, street) of 
#restaurant(s) with the most A grades.  Account for possible ties in your results.
```{r}
highest_A = sqldf(
  "SELECT dba, building, street, COUNT(grade) AS A_count
   FROM wahi_df
   GROUP BY dba
   HAVING A_count IN 
      (SELECT COUNT(grade) AS count
       FROM wahi_df
       GROUP BY dba
       ORDER BY count DESC
       LIMIT 1)
  "
)

highest_A

#The restaurant with the most A grades is the LA DINASTIA, which located at 4059 Broadway and has 17 A graede in total. 
```


#QUESTION 10 - Create a data frame called closed containing restaurants that were indicated 
#to be closed in the action field.  The data frame should contain the restaurant name, 
#address (building, street), inspection date, and action.
```{r}
action_type = sqldf(
  "SELECT action
   FROM wahi_df
   GROUP BY action
  "
)

closed = sqldf(
  "SELECT DISTINCT dba, building, street, inspection_date, action
   FROM wahi_df
   WHERE action LIKE '%Closed%'
  "
)


closed
```


#QUESTION 11 - List the restaurants included in the closed data frame and order them by number 
#of closures, from most to least.  Include the restaurant name and address.
```{r}
closure_count = sqldf(
  "SELECT dba, building, street, COUNT(inspection_date) AS close_count
   FROM closed
   GROUP BY dba, building, street
   ORDER BY close_count DESC
  "
)

closure_count
```

#QUESTION 12 - Use SQL to answer a question of your choice about restaurant violations in the 
#campus neighborhood.
```{r}
#Question: what kind of violation is most common in this neighborhood? Give the violation code and description. 

common_vio = sqldf(
  "SELECT violation_code, violation_description, COUNT(violation_code) AS vio_count
   FROM wahi_df
   GROUP BY violation_code
   HAVING vio_count IN 
      (SELECT COUNT(violation_code) AS count
       FROM wahi_df
       GROUP BY violation_code
       ORDER BY count DESC
       LIMIT 1)
  "
)

common_vio

#The most common violation is 10F.
```


