---
title: "Reading data from web"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(rvest)
library(httr)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis" 
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Scrape data from web

### First table in [National Survey on Drug Use and Health](http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm) :

Read the html: 
```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)
```

Extract the table(s): 
```{r}
tabl_mar =   
  drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>% 
  html_table() %>%  
  slice(-1) %>% #get rid of the first row
  as_tibble()
```


##Star War Movie info
Data form [IMDB Star War Page](https://www.imdb.com/list/ls070150896/):

```{r}
url = "https://www.imdb.com/list/ls070150896/"
swm_html = read_html(url)
```

[SelectorGadget](https://selectorgadget.com/) works here!!!

```{r}
title_vec = 
  swm_html %>% 
  html_nodes(css = ".lister-item-header a") %>%  # SelectorGadget 
  html_text %>% view

gross_rev_vec =  
  swm_html %>% 
  html_nodes(css = ".text-muted .ghost~ .text-muted+ span") %>%  # SelectorGadget 
  html_text %>% view

runtime_vec = 
  swm_html %>%
  html_nodes(".runtime") %>%
  html_text() %>% view

swm_df = 
  tibble(
    title = title_vec,
    gross_rev = gross_rev_vec,
    runtime = runtime_vec) %>% view
```

##NYC Water consumption data, using API 

Get data from [Water Consumption In The New York City](https://data.cityofnewyork.us/Environment/Water-Consumption-In-The-New-York-City/ia2d-e54m)

**Why API instaed of exporting:**
Sometimes data may get updating, and to make our work more reproducible. 

```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content("parsed")

#if JSON
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") %>% 
  content("text") %>%
  jsonlite::fromJSON() %>%
  as_tibble()
```


##[BRFSS](https://chronicdata.cdc.gov/Behavioral-Risk-Factors/Behavioral-Risk-Factors-Selected-Metropolitan-Area/acme-vg9e)

Same procedure for a different dataset
```{r}
brfss_2010 = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv") %>% 
  content("parsed") # only 1000 rows here, we can request more


brfss_smart2010 = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) %>% 
  content("parsed")
```

